---
title: "Sumit"
output: html_document
---
RQ5: Customers that are most profitable in terms of revenue generation

In order to find the solution to the above mentioned RQ, we have formulated loyality in terms of the ratio of products purchased and ratio of amount of money spent.

```{r echo=TRUE, include=TRUE}
library(tidyverse)
library(ggplot2)
library("RColorBrewer")


file_path<- "Input Dataset/Cleaned Dataset/Supermarket_DataCleaned.csv"
supermarket_data_clean <- read_csv(file_path)


## Ratio of products purchased from each shop (product purchased by a customer at a shop over total produccts purchased by that customer)
prod_purch_ratio_shop_1 <- with(supermarket_data_clean, products_purchased_shop_1/products_purchased_total)
prod_purch_ratio_shop_2 <- with(supermarket_data_clean, products_purchased_shop_2/products_purchased_total)
prod_purch_ratio_shop_3 <- with(supermarket_data_clean, products_purchased_shop_3/products_purchased_total)
prod_purch_ratio_shop_4 <- with(supermarket_data_clean, products_purchased_shop_4/products_purchased_total)
prod_purch_ratio_shop_5 <- with(supermarket_data_clean, products_purchased_shop_5/products_purchased_total)

## Ratio of amount spent at each shop (amount spent by a customer at a shop over total amount spent by that customer)
amt_purch_ratio_shop_1 <- with(supermarket_data_clean, amount_purchased_shop_1/amount_purchased_total)
amt_purch_ratio_shop_2 <- with(supermarket_data_clean, amount_purchased_shop_2/amount_purchased_total)
amt_purch_ratio_shop_3 <- with(supermarket_data_clean, amount_purchased_shop_3/amount_purchased_total)
amt_purch_ratio_shop_4 <- with(supermarket_data_clean, amount_purchased_shop_4/amount_purchased_total)
amt_purch_ratio_shop_5 <- with(supermarket_data_clean, amount_purchased_shop_5/amount_purchased_total)

## Loyality of customer to a shop (Ratio of products purchased from each shop times Ratio of amount spent at each shop)
loyality_shop_1 <- prod_purch_ratio_shop_1 * amt_purch_ratio_shop_1
loyality_shop_2 <- prod_purch_ratio_shop_2 * amt_purch_ratio_shop_2
loyality_shop_3 <- prod_purch_ratio_shop_3 * amt_purch_ratio_shop_3
loyality_shop_4 <- prod_purch_ratio_shop_4 * amt_purch_ratio_shop_4
loyality_shop_5 <- prod_purch_ratio_shop_5 * amt_purch_ratio_shop_5


## Data frames based on calculated values
prod_purch_ratio_shop <- data.frame(supermarket_data_clean$customer_id, prod_purch_ratio_shop_1, prod_purch_ratio_shop_2, prod_purch_ratio_shop_3, prod_purch_ratio_shop_4, prod_purch_ratio_shop_5)
amt_purch_ratio_shop <- data.frame(supermarket_data_clean$customer_id, amt_purch_ratio_shop_1, amt_purch_ratio_shop_2, amt_purch_ratio_shop_3, amt_purch_ratio_shop_4, amt_purch_ratio_shop_5)
loyality <- data.frame(supermarket_data_clean$customer_id, loyality_shop_1, loyality_shop_2, loyality_shop_3, loyality_shop_4, loyality_shop_5)

## Data frames as per shop nos
shop_1 <- data.frame(supermarket_data_clean$customer_id, prod_purch_ratio_shop_1, amt_purch_ratio_shop_1, loyality_shop_1)
shop_2 <- data.frame(supermarket_data_clean$customer_id, prod_purch_ratio_shop_2, amt_purch_ratio_shop_2, loyality_shop_2)
shop_3 <- data.frame(supermarket_data_clean$customer_id, prod_purch_ratio_shop_3, amt_purch_ratio_shop_3, loyality_shop_3)
shop_4 <- data.frame(supermarket_data_clean$customer_id, prod_purch_ratio_shop_4, amt_purch_ratio_shop_4, loyality_shop_4)
shop_5 <- data.frame(supermarket_data_clean$customer_id, prod_purch_ratio_shop_5, amt_purch_ratio_shop_5, loyality_shop_5)


## Customer segments for Shop 1 based on Loyality Score
high_end_shop_1 <- shop_1[which(shop_1$loyality_shop_1 > 0.9), ]
mid_range_shop_1 <- shop_1[which(shop_1$loyality_shop_1 > 0.50 & shop_1$loyality_shop_1 <= 0.9), ]
low_end_shop_1 <- shop_1[which(shop_1$loyality_shop_1 > 0 & shop_1$loyality_shop_1 <= 0.5), ]

## Customer segments for Shop 2 based on Loyality Score
high_end_shop_2 <- shop_2[which(shop_2$loyality_shop_2 > 0.9), ]
mid_range_shop_2 <- shop_2[which(shop_2$loyality_shop_2 > 0.50 & shop_2$loyality_shop_2 <= 0.9), ]
low_end_shop_2 <- shop_2[which(shop_2$loyality_shop_2 > 0 & shop_2$loyality_shop_2 <= 0.5), ]

## Customer segments for Shop 3 based on Loyality Score
high_end_shop_3 <- shop_3[which(shop_3$loyality_shop_3 > 0.9), ]
mid_range_shop_3 <- shop_3[which(shop_3$loyality_shop_3 > 0.50 & shop_3$loyality_shop_3 <= 0.9), ]
low_end_shop_3 <- shop_3[which(shop_3$loyality_shop_3 > 0 & shop_3$loyality_shop_3 <= 0.5), ]

## Customer segments for Shop 4 based on Loyality Score
high_end_shop_4 <- shop_4[which(shop_4$loyality_shop_4 > 0.9), ]
mid_range_shop_4 <- shop_4[which(shop_4$loyality_shop_4 > 0.50 & shop_4$loyality_shop_4 <= 0.9), ]
low_end_shop_4 <- shop_4[which(shop_4$loyality_shop_4 > 0 & shop_4$loyality_shop_4 <= 0.5), ]

## Customer segments for Shop 5 based on Loyality Score
high_end_shop_5 <- shop_5[which(shop_5$loyality_shop_5 > 0.9), ]
mid_range_shop_5 <- shop_5[which(shop_5$loyality_shop_5 > 0.50 & shop_5$loyality_shop_5 <= 0.9), ]
low_end_shop_5 <- shop_5[which(shop_5$loyality_shop_5 > 0 & shop_5$loyality_shop_5 <= 0.5), ]


## Number of customers in each segment for Shop 1
count_high_end_shop_1 <- nrow(high_end_shop_1)
count_mid_range_shop_1 <- nrow(mid_range_shop_1)
count_low_end_shop_1 <- nrow(low_end_shop_1)

## Number of customers in each segment for Shop 2
count_high_end_shop_2 <- nrow(high_end_shop_2)
count_mid_range_shop_2 <- nrow(mid_range_shop_2)
count_low_end_shop_2 <- nrow(low_end_shop_2)

## Number of customers in each segment for Shop 3
count_high_end_shop_3 <- nrow(high_end_shop_3)
count_mid_range_shop_3 <- nrow(mid_range_shop_3)
count_low_end_shop_3 <- nrow(low_end_shop_3)

## Number of customers in each segment for Shop 4
count_high_end_shop_4 <- nrow(high_end_shop_4)
count_mid_range_shop_4 <- nrow(mid_range_shop_4)
count_low_end_shop_4 <- nrow(low_end_shop_4)

## Number of customers in each segment for Shop 5
count_high_end_shop_5 <- nrow(high_end_shop_5)
count_mid_range_shop_5 <- nrow(mid_range_shop_5)
count_low_end_shop_5 <- nrow(low_end_shop_5)

## Setting up values
shops <- c(1, 1, 1, 2, 2, 2, 3, 3, 3, 4, 4, 4, 5, 5, 5)

customer_segments <- c("high", "med", "low", "high", "med", "low", "high", "med", "low", "high", "med", "low", "high", "med", "low")

count_shop_1 <- c(count_high_end_shop_1, count_mid_range_shop_1, count_low_end_shop_1)
count_shop_2 <- c(count_high_end_shop_2, count_mid_range_shop_2, count_low_end_shop_2)
count_shop_3 <- c(count_high_end_shop_3, count_mid_range_shop_3, count_low_end_shop_3)
count_shop_4 <- c(count_high_end_shop_4, count_mid_range_shop_4, count_low_end_shop_4)
count_shop_5 <- c(count_high_end_shop_5, count_mid_range_shop_5, count_low_end_shop_5)

counts <- c(count_shop_1, count_shop_2, count_shop_3, count_shop_4, count_shop_5)


## Count of customers based on the three customer segments (high, mid, low)
shop_customer_segment_count <-data.frame(shops, customer_segments, counts)

## Change datatype of columns to numerical
shop_customer_segment_count <- shop_customer_segment_count %>% mutate_if(is.integer, as.numeric) %>% mutate_if(is.factor, as.numeric)
```

This graph plots the count of customers for each shop based on three different customer segments (high, mid, low)
```{r}
## Plot the count of customers for each shop based on customer segments (high - 1, mid - 3, low - 2)
qplot(shops, counts, data = shop_customer_segment_count, color = customer_segments)
qplot(shops, counts, data = shop_customer_segment_count, color = customer_segments, facets = .~ customer_segments)
```

This graph plots the count of customers for each cutomer segment (high, mid, low)
```{r}
## Plot the count of customers for each segment
qplot(customer_segments, counts, data = shop_customer_segment_count)
```

