---
title: "RQ3"
author: "Sumit Kundu"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Wrangling

We removed a few unimportant and derieved features while preparing the dataset for RQ1. We repeated the same here as well. We also generated the most preferred shop (e) for each customer based on the same assumption we used for RQ1 and used it as the class label.

We generated new columns to find the average distance travelled (a), average number of products purchased (b), average number of unique products purchased (c), average product price (d) and average amount spent (e) for each customer. We further generated a dataset using these newly created columns - (a), (b), (c), (d) and (e) to be used for creating a model that would predict the shop most likely to be selected by a customer as their most preferred shop.

```{r echo=TRUE, message=FALSE}
## Call the packages
library(tidyverse)
library(dplyr)

## Set file path
file_path <- "/home/sumit/Data-Science-with-R/Input Dataset/Cleaned Dataset/Supermarket_DataCleaned.csv"

## Read data from a file
supermarket_data_clean <- read_csv(file_path)

## Generate columns to find the average distance travelled, average number of products purchased, average number of unique products purchased, average product price
## and average amount spent by each customer
supermarket_data_clean$distance_avg <- with(supermarket_data_clean, avg_distance_to_all_shops)
supermarket_data_clean$products_purchased_avg <- with(supermarket_data_clean, round(products_purchased_total/shops_used_total))
supermarket_data_clean$unique_products_purchased_avg <- with(supermarket_data_clean, round(unique_products_purchased_total_exclCommon/shops_used_total))
supermarket_data_clean$product_price_avg<- with(supermarket_data_clean, avg_purchased_product_price_allShops)
supermarket_data_clean$amount_purchased_avg<- with(supermarket_data_clean, avg_purchase_amount_allShops)

## Generate new columns to find shops from which most products are purchased, shops from which most unique products are purchased, 
## shops offfering least average product price and shops at which maximum amount of money is spent by a customers.  
supermarket_data_clean$most_prod_purch_from <- str_sub(names(supermarket_data_clean[,11:15]),-6,-1)[max.col(supermarket_data_clean[,11:15], "random")]
supermarket_data_clean$most_uni_prod_purch_from <- str_sub(names(supermarket_data_clean[,17:21]),-6,-1)[max.col(supermarket_data_clean[,17:21], "random")]
supermarket_data_clean$least_avg_prod_pri <- str_sub(colnames(supermarket_data_clean[,23:27]),-6,-1)[apply(supermarket_data_clean[,23:27],1,which.min)]
supermarket_data_clean$max_amt_purch <- str_sub(names(supermarket_data_clean[,29:33]),-6,-1)[max.col(supermarket_data_clean[,29:33], "random")]

## Create a data frame having only the newly generated columns
gen_df <- supermarket_data_clean[,46:49]

## Gnereate new columns - most preferred shop and class to which the customer belongs
for (row in 1:nrow(gen_df)){
  ## Create a vector for each row
  vec <- c(gen_df[row, "most_prod_purch_from"], gen_df[row, "most_uni_prod_purch_from"], gen_df[row, "max_amt_purch"])
  
  ## Sort and find the most preferred shop
  supermarket_data_clean[row, "most_pref_shop"] <- names(sort(summary(as.factor(unlist(vec))), decreasing=T)[1:1])
}

## Re-order columns
supermarket_data_predict <- supermarket_data_clean[c(41,42,43,44,45,50)]

## Set file path 
clean_filepath = "~/Data-Science-with-R/Input Dataset/Cleaned Dataset/Supermarket_Data_Prediction.csv"

## Write the dataframe to csv file
write.csv(supermarket_data_predict, file = clean_filepath, row.names = FALSE)

```